---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r loadingData, echo=TRUE}
library(dplyr)
library(lubridate)

if(!file.exists("activity.csv")){
        unzip("activity.zip")
}
activity <- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?
```{r totalSteps}
daily_counts <-  activity %>%
        mutate(date_by_day = date(date)) %>%
        group_by(date_by_day) %>%
        summarize(steps = sum(steps, na.rm = TRUE))
```
Histogram for steps per day
```{r histogramOfSteps}
hist(daily_counts$steps, xlab = "Number of steps")
```

The mean of the daily steps is
```{r meanSteps}
mean(daily_counts$steps)
```

The median of the daily steps is
```{r medianSteps}
median(daily_counts$steps)
```

## What is the average daily activity pattern?
```{r averageSteps}
daily_average <-  activity %>%
        mutate(date_by_day = date(date)) %>%
        group_by(date_by_day) %>%
        summarize(avgSteps = mean(steps, na.rm = TRUE))
daily_average <- daily_average[!is.na(daily_average$avgSteps),]
plot(daily_average$date_by_day, daily_average$avgSteps, main = "Average steps by day", xlab = "day", ylab = "average steps", type = "l")
abline(v=daily_average[daily_average$avgSteps == max(daily_average$avgSteps, na.rm = TRUE),"date_by_day"], col = "red")
```

## Imputing missing values
The total missing values in the dataset are:
```{r naValues}
sum(is.na(activity$steps))
```
Filling the missing values
```{r fillingValues}
activity2 <- activity[, c(1, 2)]
activity2$steps <- ifelse(is.na(activity2$steps), 0, activity2$steps)
head(activity2)
```

Calculating the total steps per day and compare with the mean and the median
```{r histogramData}
daily_counts2 <-  activity2 %>%
        mutate(date_by_day = date(date)) %>%
        group_by(date_by_day) %>%
        summarize(steps = sum(steps, na.rm = TRUE))

daily_average2 <-  activity2 %>%
        mutate(date_by_day = date(date)) %>%
        group_by(date_by_day) %>%
        summarize(avgSteps = mean(steps, na.rm = TRUE))

daily_median2 <-  activity2 %>%
        mutate(date_by_day = date(date)) %>%
        group_by(date_by_day) %>%
        summarize(medianSteps = median(steps, na.rm = TRUE))
par(mfrow = c(1,3))
hist(daily_counts2$steps, col = "red")
hist(daily_average2$avgSteps, col = "blue")
hist(daily_median2$medianSteps, col = "green")
```


## Are there differences in activity patterns between weekdays and weekends?
```{r weekdays}
activity2$weekday <- ifelse(weekdays(date(activity2$date))=="Monday" | weekdays(date(activity2$date))=="Tuesday" | weekdays(date(activity2$date))=="Wednesday" | weekdays(date(activity2$date))=="Thursday" | weekdays(date(activity2$date))=="Friday", "weekday", "weekend")

average_weekday <- activity2 %>%
        group_by(weekday) %>%
        summarize(avgSteps = mean(steps, na.rm = TRUE))

plot(date(activity2$date), activity2$steps, type = "l")
abline(h=average_weekday$avgSteps[1], col = "red")
abline(h=average_weekday$avgSteps[2], col = "blue")

```